\bf Are you normal? A new projection pursuit index to assess a sample against a multivariate null distribution

### 2024-06-25 paper.qmd
sample code in implementation

This was implemented in the `tourr` (REF) package, where the projected ellipsoid can be drawn for each projection.

>>>>

This is implemented in the `tourr` (REF) package, where the projected ellipsoid can be drawn for each projection. The guided tour will take arguments specifying the data, and the null variance-covariance matrix.

```{r}
#| eval: false
#| echo: true
library(tourr)
library(mulgar)
set.seed(929)
vc_null <- matrix(rep(0.5, 5*5), ncol=5)
diag(vc_null) <- 1
m_null <- rep(0, 5)
vc_samp <- matrix(rep(0, 5*5), ncol=5)
diag(vc_samp) <- 1
vc_samp[4,5] <- -0.47
vc_samp[5,4] <- -0.56
vc_samp <- vc_samp*0.1
m_samp <- c(0, 0, 0, 1.9, 2.3)
samp <- as.data.frame(rmvn(6, 
                           mn = m_samp, 
                           vc = vc_samp))
animate_xy(samp, guided_anomaly_tour(anomaly_index(),
  ellipse=vc_null), ellipse=vc_null, 
  axes = "bottomleft", half_range=5, center=FALSE)
```

::: {#fig-anomaly layout-ncol=2}

![Random projection.](images/anomaly1.png){#fig-anomaly1 width=300 align="center"}

![Optimal projection.](images/anomaly2.png){#fig-anomaly2 width=300 align="center"}

Two projections of simulated example data corresponding to the sample code: (a) Random projection where sample is inside the 2-D ellipse, (b) Optimal projection from index, showing most of the sample outside. A red cross indicates that the point is outside the p-D ellipse. The optimal projection uses mostly variables $x_4, x_5$, which is expected because these are the two directions where the sample most differs from the norm.
:::

### 2024-06-25 paper.tex
sample code in implementation

@@

This was implemented in the \texttt{tourr} (REF) package, where the
projected ellipsoid can be drawn for each projection.

>>>>

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

@@

This is implemented in the \texttt{tourr} (REF) package, where the
projected ellipsoid can be drawn for each projection. The guided tour
will take arguments specifying the data, and the null
variance-covariance matrix.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tourr)}
\FunctionTok{library}\NormalTok{(mulgar)}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{929}\NormalTok{)}
\NormalTok{vc\_null }\OtherTok{\textless{}{-}} \FunctionTok{matrix}\NormalTok{(}\FunctionTok{rep}\NormalTok{(}\FloatTok{0.5}\NormalTok{, }\DecValTok{5}\SpecialCharTok{*}\DecValTok{5}\NormalTok{), }\AttributeTok{ncol=}\DecValTok{5}\NormalTok{)}
\FunctionTok{diag}\NormalTok{(vc\_null) }\OtherTok{\textless{}{-}} \DecValTok{1}
\NormalTok{m\_null }\OtherTok{\textless{}{-}} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{5}\NormalTok{)}
\NormalTok{vc\_samp }\OtherTok{\textless{}{-}} \FunctionTok{matrix}\NormalTok{(}\FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{5}\SpecialCharTok{*}\DecValTok{5}\NormalTok{), }\AttributeTok{ncol=}\DecValTok{5}\NormalTok{)}
\FunctionTok{diag}\NormalTok{(vc\_samp) }\OtherTok{\textless{}{-}} \DecValTok{1}
\NormalTok{vc\_samp[}\DecValTok{4}\NormalTok{,}\DecValTok{5}\NormalTok{] }\OtherTok{\textless{}{-}} \SpecialCharTok{{-}}\FloatTok{0.47}
\NormalTok{vc\_samp[}\DecValTok{5}\NormalTok{,}\DecValTok{4}\NormalTok{] }\OtherTok{\textless{}{-}} \SpecialCharTok{{-}}\FloatTok{0.56}
\NormalTok{vc\_samp }\OtherTok{\textless{}{-}}\NormalTok{ vc\_samp}\SpecialCharTok{*}\FloatTok{0.1}
\NormalTok{m\_samp }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\FloatTok{1.9}\NormalTok{, }\FloatTok{2.3}\NormalTok{)}
\NormalTok{samp }\OtherTok{\textless{}{-}} \FunctionTok{as.data.frame}\NormalTok{(}\FunctionTok{rmvn}\NormalTok{(}\DecValTok{6}\NormalTok{, }
                           \AttributeTok{mn =}\NormalTok{ m\_samp, }
                           \AttributeTok{vc =}\NormalTok{ vc\_samp))}
\FunctionTok{animate\_xy}\NormalTok{(samp, }\FunctionTok{guided\_anomaly\_tour}\NormalTok{(}\FunctionTok{anomaly\_index}\NormalTok{(),}
  \AttributeTok{ellipse=}\NormalTok{vc\_null), }\AttributeTok{ellipse=}\NormalTok{vc\_null, }
  \AttributeTok{axes =} \StringTok{"bottomleft"}\NormalTok{, }\AttributeTok{half\_range=}\DecValTok{5}\NormalTok{, }\AttributeTok{center=}\ConstantTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{minipage}{0.50\linewidth}

\includegraphics[width=3.125in,height=\textheight]{images/anomaly1.png}

\subcaption{\label{fig-anomaly1}Random projection.}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\includegraphics[width=3.125in,height=\textheight]{images/anomaly2.png}

\subcaption{\label{fig-anomaly2}Optimal projection.}

\end{minipage}%

\caption{\label{fig-anomaly}Two projections of simulated example data
corresponding to the sample code: (a) Random projection where sample is
inside the 2-D ellipse, (b) Optimal projection from index, showing most
of the sample outside. A red cross indicates that the point is outside
the p-D ellipse. The optimal projection uses mostly variables
\(x_4, x_5\), which is expected because these are the two directions
where the sample most differs from the norm.}
